---
title: "Hands-on_Ex04"
author: "Lucas Luo"
date: "May 07 2025"
date-modified: "last-modified"

format:
  html:
    code-tools: true          
    code-fold: true          
    toc: true
    theme: lux

execute: 
  eval: true
  echo: true
  message: false
  warning: false
  freeze: true
---

# **Fundamentals of Visualization Analytics**

# **1 Visualizing Distribution**

The following section was modified according to https://r4va.netlify.app/chap09.

```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)

exam_data <- read_csv("data/Exam_data.csv")
```

Visualizing Distribution with Ridgeline Plot When to Use Ridgeline Plots:

When you have a medium to high number of groups (more than 5), as it saves space. If there is a clear pattern or ranking among groups, making insights easier to spot. When you want to efficiently visualize distributions without using separate plots. When Not to Use Ridgeline Plots:

If there are fewer than 5 groups, other distribution plots (e.g., boxplots) may be better. When groups lack a clear pattern, as overlapping can make the plot messy. If hiding some data due to overlap reduces clarity rather than improving visualization.

```{r}
ggplot(data = exam_data, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("gray", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, 
                   expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### **1.2 Varying fill colors along the x axis**
